### 2024-07-09 base-rates.qmd
20240608 - getting started
20240627 - sessionInfo() for reproducibility
20240627 - sessionInfo()
20240703 - conclusion section
20240708 - base rate fallacy

## Getting Started

### Load Packages

@@

>>>>

## Getting Started {#sec-baseRatesGettingStarted}

### Load Packages {#sec-baseRatesLoadPackages}

@@

## Overview {#sec-baseRatesConclusion}

The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For instance, the base rate of HIV is the probability of developing HIV.
In the U.S., [the prevalence rate of HIV is ~0.4% of the adult population](https://map.aidsvu.org/profiles/nation/usa/overview) (archived at <https://perma.cc/8GE6-GAPC>).

## Conclusion {#sec-baseRatesConclusion}

::: {.content-visible when-format="html"}

## Session Info {#sec-baseRatesSessionInfo}

```{r}
sessionInfo()
```

:::

### 2024-07-09 cognitive-bias.qmd
20240708 - base rate fallacy

In our everyday lives, humans are presented with overwhelming amounts of information.
Because human minds cannot parse every piece of information equally, we tend to take mental shortcuts, called *heuristics*.
These mental shortcuts can be helpful.
They reduce our mental load and can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
However, these mental shortcuts can also lead us astray and make systematic errors in our judgments and predictions.
*Cognitive biases* are systematic errors in thinking.

@@

@@

@@

@@

- base rate neglect
- forgetting about regression to the mean
- overconfidence
- confirmation bias
- hindsight bias
- loss aversion bias
- endowment bias
- Dunning–Kruger effect

@@

- sunk cost fallacy
- hot hand fallacy
- gambler's fallacy

>>>>

In our everyday lives, humans are presented with overwhelming amounts of information.
Because human minds cannot parse every piece of information equally, we tend to take mental shortcuts, called *heuristics*.
These mental shortcuts can be helpful.
They reduce our mental load and can help us make quick judgments to stay alive or to make complex decisions in the face of uncertainty.
However, these mental shortcuts can also lead us astray and to make systematic errors in our judgments and predictions.
*Cognitive biases* are systematic errors in thinking.
*Fallacies* are forms of flawed reasoning.

@@

### Availability Heuristic {#sec-heuristicsAvailability}

@@

### Representativeness Heuristic {#sec-heuristicsRepresentativeness}

@@

### Anchoring and Adjustment Heuristic {#sec-heuristicsAnchoringAdjustment}

@@

- overconfidence bias
- confirmation bias
- recency bias
- hindsight bias
- loss aversion bias
- endowment bias
- bandwagon effect bias
- Dunning–Kruger effect bias

### Overconfidence Bias {#sec-cognitiveBiasesOverconfidence}

### Confirmation Bias {#sec-cognitiveBiasesConfirmation}

### Recency Bias {#sec-cognitiveBiasesRecency}

### Hindsight Bias {#sec-cognitiveBiasesHindsight}

### Loss Aversion Bias {#sec-cognitiveBiasesLossAversion}

### Endowment Bias {#sec-cognitiveBiasesEndowment}

### Bandwagon Effect Bias {#sec-cognitiveBiasesBandwagon}

### Dunning–Kruger Effect Bias {#sec-cognitiveBiasesDunningKruger}

@@

- base rate fallacy (aka base rate neglect)
- regression fallacy
- sunk cost fallacy
- hot hand fallacy
- gambler's fallacy

### Base Rate Fallacy {#sec-fallaciesBaseRate}

The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
For example, the base rate of work-related injury is the general probability of experiencing a work-related injury, irrespective of other factors (e.g., the type of job, the person's age, the person's sex).
Among the working population in the U.S., the lifetime prevalence of work-related injuries (i.e., the percent of people who will experience a work-related injury at some point in their lives), is ~35% (<https://www.cdc.gov/mmwr/volumes/69/wr/mm6913a1.htm>; archived at <https://perma.cc/A2L6-WPEH>).
Thus, the base rate of work-related injuries in the U.S. is ~35%.
The probability of work-related injuries is higher for some occupations (e.g., construction) and for some groups (e.g., men, 55–64-year-olds, Black or Multiracial, who are self-employed and have less than high school education) than others.
Nevertheless, is we ignore all of the interacting factors, the general probability of work-related injuries is 35%.
If we made a prediction that someone would be highly likely (> 90%) to experience a work-related injury because they are male and self-employed, this would be ignoring the relatively lower base rate of work-related injury.
Indeed, even men (36.7%) and self-employed individuals (41.2%) have less than a 50% chance of experiencing a work-related injury.

As applied to fantasy football, consider that you read about a potential sleeper Wide Receiver who had a stellar performance in a preseason game.
If you select this player early on in the draft based on this information, this would be ignoring the general probability that most players who have a strong performance in the preseason do not perform as well in the regular season.
Performance in the preseason is not strongly predictive of performance in the regular season (<https://fivethirtyeight.com/features/the-nfl-preseason-is-not-predictive-but-it-can-often-seem-that-way/>; archived at <https://perma.cc/FSG2-6AXE>).

### Regression Fallacy {#sec-fallaciesRegression}

### Sunk Cost Fallacy {#sec-fallaciesSunkCost}

### Hot Hand Fallacy {#sec-fallaciesHotHand}

### Gambler's Fallacy {#sec-fallaciesGambler}
